Web Processing - Standardised GIS Analyses for Cable Route Planning

### 2023-03-15 paper/discussion.tex
send paper

% !TeX spellcheck = en_GB
In theory, the need for computing time increases with the resolution as power of two. 
Similarly, the use of main memory increases. 
This again limits the number data points, that can be processed and resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls has been used for the aggregation of higher resolution.

@@

For b), two methods have been used to increase the similarity of the paths computed from the medium resolution raster, to the path of the highest resolution raster.
These methods are used as surrogates for the more complex calculation of the Least Cost Path with the higher resolution.
In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all touched set to False and True raster were averaged in different ratios, to compute the optimum weighted cost raster.
While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
This disadvantage could be reduced by normalising the costs.

This shows, that the way the cost raster is created in the first place can play a crucial role, in the end result.
So that a nuance, can cause a detour.
When this behaviour occurs, the polygon may not include the Least Cost Path.
This polygon should therefore be overlapped with a polygon around the shortest path.
For the set of control paths downsampling could not outperform the original medium resolution rasters.

@@

The set of rules, that are used to create the intermediate cost raster, includes a rule to create buffer around buildings which is set to the level \textit{prohibited} areas.
In all touched set to False rasters, the resolution of the medium level raster needs to be high enough to show every detail.
At least in the magnitude of the minimum object size plus twice its buffer.
This is true for the 10~m resolution raster and less true for the 25~m resolution raster, that misses some details for roads for all touched False raster.
Other details such as rivers and houses, are already included in the lower resolution raster, due to larger buffers.
The raster with all~touched set to False might miss some objects, but the missing chance is propotional to the object size and the extent of overestimation is reduced.
The Least Cost Path algorithm searches for an optimal path as a line.
As lines do not have a width, the route found might contain bottlenecks, that have a smaller width than the object that should be placed there.
Therefore, the used resolution should not be smaller than, the width of the object that should be placed.
This can not be avoided by downsampling, but by weighting the medium resolution all~touched True and False rasters.

This paper examines the effect of computational costs and deviation of the results, is examined for a very limited set of points.
Also, only the cost of finding the Least Cost Path from a single start point, to a single end point has been considered.
If multiple endpoints are used, the computational cost for the aggregated cost raster has only to be paid once.

If multiple paths are calculated from a single raster, the speed-up benefit is reduced.
Especially, pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective, because as the number of paths increases, fewer pixels are clipped.
The Least Cost Path algorithm does only select the single most cost-effective path.
Therefore, paths of similar, but slightly higher costs remain unknown.
In additional, slight variations on the costs rasters can lead to very different paths, although the costs will not change much.
An end-user may be interested in selecting a path from a set of similar aggregated costs and applying their own evaluation criteria.
This can be achieved a by adjusting the backtracking and return polygons, or by applying perturbation on the costs.

In this work, the intermediate cost raster layer are aggregated using the maximum function. 
Other possible aggregation functions are the sum and average functions.
Each aggregation function can be justified, by a different interpretation of the cost and its scale.

When the \textit{Prohibited} level is used as the highest level, then summing the two highest levels would result in a new highest level. 
Also, the maximum function does not interfere with the nodata value. 
This can be done by a nansum- / nanmean-function, if the nodata value is set to \textit{not a number} during the aggregation.
The disadvantage of aggregation with the maximum function is, that this aggregation is unable to distinguish between nuances of different overlapping intermediate costs.
On sum or average aggregated rasters, one can distinguish between, different sublevels.

The fact that all touched False rasters produce more similar results than high resolution raster, is probably due to the fact, that the default level is relatively low.
As the default level increases, the effect would probably be reduced for low resolution rasters.
For low resolution raster, the effect would still be present, because the fact, that the raster pixel centre is used for sampling, reflects the original geometry better.

@@

The all~touched set to True cost raster shows every detail, but the sampling with all~touched set to True increases the size of the features.
The fact that the aggregated costs per resolution for all touched True rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, because the costs are exponential scaled.

>>>>

% !TeX spellcheck = en_GB
In theory, the need for computing time increases with the resolution as power of two. 
Similarly, the use of main memory increases. 
This again limits the number of data points, that can be processed and the resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls has been used for the aggregation of higher resolution.

@@

For b), two methods have been used to increase the similarity of the paths computed from the medium resolution raster, to the paths of the highest resolution raster.
These methods are used as surrogates for the more complex calculation of the Least Cost Path with the higher resolution.
In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all~touched set to False and True rasters were averaged in different ratios, to compute the optimum weighted cost raster.
While the second method of using an averaged raster, shows a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
This disadvantage could be reduced by normalising the costs.
The fact that downsampling leads to paths that are closer to all~touched True, can be attributed to the fact, that all objects are included, such as in to all~touched True.

This shows that the way the cost raster is created in the first place can play a crucial role in the end result.
So, a nuance can cause a detour.
When this behaviour occurs, the polygon may not include the Least Cost Path.
This polygon should therefore be overlapped with a polygon around the shortest path.
For the set of control paths downsampling could not outperform the original medium resolution rasters.

@@

The set of rules, that are used to create the intermediate cost raster, includes a rule to create buffer around buildings which is set to the level \textit{Prohibited} areas.
In all touched set to False rasters the resolution of the medium level raster needs to be high enough to show every detail,
at least in the magnitude of the minimum object size plus twice its buffer.
This is true for the 10~m resolution raster and less true for the 25~m resolution raster, that misses some details for roads for all touched False raster.
Other details such as rivers and houses are already included in the lower resolution raster, due to larger buffers.
The raster with all~touched set to False might miss some objects, but the missing chance is propotional to the object size and the extent of overestimation is reduced.
The Least Cost Path algorithm searches for an optimal path as a line.
As lines do not have a width, the route found might contain bottlenecks, that have a smaller width than the object that should be placed there.
Therefore, the used resolution should not be smaller than the width of the object that should be placed.
This can not be avoided by downsampling, but by weighting the medium resolution all~touched True and False rasters.

This paper examines the effect of computational costs and deviation of the results for a limited set of points.
Also, only the cost of finding the Least Cost Path from a single starting point to a single end point has been considered.
If multiple endpoints are used, the computational cost for the aggregated cost raster has to be paid only once.

If multiple paths are calculated from a single raster, the speed-up benefit is reduced.
Especially pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective, because as the number of paths increases, fewer pixels are clipped.
The Least Cost Path algorithm does only select the single most cost-effective path.
Therefore, paths of similar, but slightly higher costs remain unknown.
In additional, slight variations on the costs rasters can lead to very different paths, although the costs will not change much.
End-users may be interested in selecting a path from a set of similar aggregated costs and applying their own evaluation criteria.
This can be achieved a by adjusting the backtracking and return polygons, or by applying perturbation on the costs.

In this work, the intermediate cost layers are aggregated by the maximum function. 
Other possible aggregation functions are the sum and average functions.
Each aggregation function can be justified by a different interpretation of the cost and its scale.

When the \textit{Prohibited} level is used as the highest level, then summing the two highest levels would result in a new highest level. 
Also, the maximum function does not interfere with the nodata value. 
This can be done by a nansum- / nanmean-function, if the nodata value is set to \textit{not a number} during the aggregation.
The disadvantage of aggregation with the maximum function is, that this aggregation is unable to distinguish between nuances of different overlapping intermediate costs.
On sum or average aggregated rasters, one can distinguish between different sublevels.

All touched False rasters produce more similar results than all touched True raster, which is probably due to the fact, that the default level is relatively low.
As the default level increases, the effect would probably be reduced for low resolution rasters.
For high resolution raster, the effect would still be present, because the use of the pixel centre for sampling, reflects the original geometry better.

@@

The all~touched set to True cost raster shows every detail, but the sampling with all~touched set to True increases the size of the features.
The fact that the aggregated costs per resolution for all touched True rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, as they are exponential scaled.

### 2023-03-15 paper/methods.tex
send paper

The Least Cost Path algorithm is used to plan a possible route for a power line.
The Least Cost Path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
This makes it possible to find routes on graphs that are not predefined, such as road networks.
The weights of the edges are the local costs of transit from one pixel center to the neighbouring pixel center.
The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate to transverse a given land usage.
The Least Cost Path algorithm consists of at least two successive steps.
1) The first step is to aggregate the costs of travelling from the starting point to a given set of end points.
This step generates the aggregated cost raster of travelling from the starting point to any point of the cost raster.
2) In the next step the backtracking, the route of the actual Least Cost Path is calculated.
For each end point the path via the lowest cost neighbour is taken until the starting point is reached.

@@

PyWPS~\cite{noauthor_welcome_2016} is used to provide the Least Cost Path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster and receives the resulting Least Cost Path.
The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0, but refactored to optionally export the aggregated costs.

@@

All these layers are provided as vectors.
The Least Cost Path algorithm uses raster data.
Rasterisation transforms a vector into a raster.
Rasterisation can be done in two different ways.
In both ways, the rasterisation can be imaged, as the old vector is superimposed on the new raster grid with the new
given resolution and the new affine transformation and the coordinate reference system of the vector.
Both rasterisation techniques differ in the selection of the pixel, that describe the original polygon.
A pixels will be selected, if either the centre of that pixel is overlapped
by the geo-object, or any part of the pixel is overlaid.
Setting all~touched to True implies the version with any part of the pixel selected.
The version, where an overlapped pixel centre is required, is setting the parameter all~touched to False.
All~touched set to False is considered the default (see figure~\ref{fig:alltouched}).
\begin{figure}[!ht]
	\centering

@@

\begin{table}[h!]
	\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
	\label{tab:1}
	\centering
	\begin{tabular}{ l  r  l }
		Cost Level 			& Cost 					& Example\\
		\hline
		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
		Strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
		No Restriction 		& 0.5					& Default\\
		Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
	\end{tabular}
\end{table}

>>>>

The Least Cost Path algorithm is used to plan a potential, cost-optimised route for a power line.
The Least Cost Path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
This makes it possible to find routes on graphs that are not predefined, such as road networks.
The weights of the edges are the local costs of transit from one pixel center to the neighbouring pixel center.
The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate to transverse a given land usage.
The Least Cost Path algorithm consists of at least two successive steps.
1) The first step is to aggregate the costs of travelling from the starting point to a given set of end points.
This step generates the aggregated cost raster of travelling from the starting point to any point of the cost raster.
2) In the next step the backtracking, the route of the actual Least Cost Path is calculated.
For each end point the path via the lowest cost neighbour is taken until the starting point is reached.

@@

PyWPS~\cite{noauthor_welcome_2016} is used to provide the Least Cost Path algorithm as a \acrshort{wps} in combination with flask~\cite{noauthor_flask_nodate}.
As client, Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster, starting point, end points and receives the resulting Least Cost Path.
The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0, but refactored to optionally export the aggregated costs in a command line tool.
On top, the \acrshort{wps} provides the complete Least Cost Path algorithm as a single capability.

@@

All these layers are provided as vectors.
The Least Cost Path algorithm uses raster data.
Rasterisation transforms a vector into a raster.
Rasterisation can be done in two different ways.
In both ways, the rasterisation can be imaged, as the old vector is superimposed on the new raster grid with the new
given resolution and the new affine transformation and the coordinate reference system of the vector.
Both rasterisation techniques differ in the selection of the pixel, that describe the original polygon.
A pixels will be selected, if either the centre of that pixel is overlapped
by the geo-object, or any part of the pixel is overlaid.
Setting all~touched to True implies the version with any part of the pixel selected.
The version, where an overlapped pixel centre is required, is setting the parameter all~touched to False.
All~touched set to False is considered the default (see figure~\ref{fig:alltouched}).
\begin{figure}[!t]
	\centering

@@

\begin{table}[t!]
	\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
	\label{tab:1}
	\centering
	\begin{tabular}{ l  r  l }
		Cost Level 			& Cost 					& Example\\
		\hline
		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
		Strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
		No Restriction 		& 0.5					& Default\\
		Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
	\end{tabular}
\end{table}

### 2023-03-15 paper/paper-CableRoute.tex
send paper

\documentclass[acmtog]{acmart}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[acronym]{glossaries}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{makecell}

\AtBeginDocument{%
	\providecommand\BibTeX{{%
			Bib\TeX}}}

\setcopyright{acmcopyright}
\copyrightyear{2022}
\acmYear{2022}
\acmDOI{XXXXXXX.XXXXXXX}

@@

	\begin{abstract}
		Least Cost Paths for a power line from a power station at the shore line to a power station further inland are calculated.
		Downsampling and clipping are investigated as techniques to speed up the calculation.
		The paths from the clipped cost raster often are the same as the original paths.
		The speed-up of the clipping can not be known in advance.
	\end{abstract}

@@

	\received{14 March 2023}
	%%\received[revised]{12 March 2009}
	%%\received[accepted]{5 June 2009}

@@

	\section{Introduction}\label{sec:introduction}

@@

@@

	\section{Conclusion}\label{sec:conclusion}
	The cost of computing of the Least Cost Path scales with the square of the resolution.
	The difference between the aggregated costs of the paths per resolution computed from the all~touched set to False raster True raster only shows a linear decrease.
	Therefore, the gain in accuracy per compute time decreases.
	The presented methods attempt to circumvent this bottleneck.
	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the Least Cost Path from the high resolution raster.
	The paths can change significantly even for a small change in total cost.
	Therefore, to develop an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy, to offer alternative paths and at the same time show variations between these paths.
	An alternative would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster. 
	This again increases the computation time.

	In an actual search for Least Cost Path search, a survey could be used as a method to estimate the costs.
	If the sample size of the survey is large enough, the
	weights could take  into account local differences in relative acceptance.

>>>>

\documentclass[acmtog]{acmart}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[acronym]{glossaries}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{makecell}
\usepackage{hyperref}

\makeatletter
\AtBeginDocument{%
	\hypersetup{
		pdftitle = {\@title},
		pdfauthor = {\@author}
	}
	\providecommand\BibTeX{{%
			Bib\TeX}}}
\makeatother

\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

@@

	\begin{abstract}
		A lot of wind energy is produced in the north of Germany. The power has to be transported further inland. To reduce congestion in the power system, new power lines have to be installed.
		A potential path for a power line from the northern shore is planned using the Least Cost Path algorithm.
		The planning is implemented in Python and provided via a \acrfull{wps} with PyWPS~\cite{noauthor_welcome_2016} and flask~\cite{noauthor_flask_nodate}.
		Downsampling and clipping are investigated as techniques to speed up the calculation.
		The paths from the clipped cost raster often are the same as the original paths.
		The speed-up of the clipping can not be known in advance.
	\end{abstract}

@@

	\received{15 March 2023}
	%%\received[revised]{12 March 2009}
	%%\received[accepted]{5 June 2009}

@@

	\section{Introduction and Objectives}\label{sec:introduction}

@@

	This paper aims to find potential paths for power lines by using the Least Cost Path algorithm. Land usage and planning data are used to estimate the costs arising by using the local area. The task was to provide the Least Cost Path using a \acrfull{wps}. 
	Beyond that, methods to reduce the needed compute power for finding the Least Cost Path had to be studied.

@@

	\section{Conclusion}\label{sec:conclusion}
	The computing costs of the Least Cost Path scales with the square of the resolution.
	The difference between the aggregated costs of the paths per resolution computed from the all~touched set to False raster and the all~touched set to True raster only shows a linear decrease.
	Therefore, the gain in accuracy per compute time decreases.
	The presented methods attempt to circumvent this bottleneck.
	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced without compromising the run time for worse distances to the Least Cost Path from the high resolution raster.
	The paths can change significantly even for a small change in total cost.
	Therefore, developing an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy to offer alternative paths and at the same time show variations between these paths.
	An alternative would be to specify a range of costs and compute the aggregation of the costs and than superimpose these aggregations and compute the path by backtracking from the superimposed cost raster. 
	This again increases the computation time.

	In an actual search for the Least Cost Path of a power, a survey could be used as a method to estimate the costs.
	If the sample size of the survey is large enough, the
	weights could take into account local differences in relative acceptance.

### 2023-03-15 paper/results.tex
send paper

\begin{table*}[t]
	\caption{Least Cost Paths as length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$) and the maximum minimum distance~($d_{max}$) and the agg. costs. From the agg. costs the differences of the agg. costs and the agg. costs per resolution are computed.} 
	\label{tab:2}
	\centering
	\begin{tabular}{ r  r  r  r  r  r  r  r  r  r}
		res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m & $d_{max}/m$ & agg.  $ cost_{al=f}$ & agg. $ cost_{al=t}$ &  $\Delta $ costs & agg. $costs_{al=f} /m$ & agg. $costs_{al=t} /m$ \\
		\hline
		5 	& 76136.3	& 78002.0 &  126.0 & 1065.0 & 18665.9 & 19616.8 & -850.00 & 93329.6 &  97584.8 \\
		10 	& 75430.1 	& 77936.6 &  277.9 & 1590.0 &  8931.2 &  9731.2 & -799.95 & 89312.5 &  97311.8 \\
		25 	& 75422.9 	& 78422.9 &  313.8 & 1621.2 &  3354.9 &  3872.7 & -517.78 & 83871.7 &  96816.4 \\
		50 	& 76135.0	& 70620.0 & 1140.0 & 4950.0 &  1409.0 &  2300.1 & -891.05 & 70451.2 & 115003.7 \\
		100 & 76283.8	& 74120.7 & 1946.4 & 6016.6 &   640.5 &  1572.3 & -931.70 & 64051.6 & 167226.8 \\

@@

As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all~touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (bi-linear) interpolation.

The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all~touched False) shows, that only downsampling to a resolution of 10~m produces a path that is relatively close the high resolution paths(see table~\ref{tab:5}).

@@

Bi-linear downsampling the of the high resolution raster to a medium resolution did not result in any benefits compared to an original medium resolution raster.
The aggregated cost per resolution of the Least Cost Path from the downsampled raster is higher than the normal medium resolution 10~m raster.
In addition, the distance from these paths to the high resolution path  is greater than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
\todo{interpretation: Why is the downsampling result so different? Don't know.}

\begin{table}[h!]
	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched False raster for the four control routes, for the reference path constructed from the 5~m  and 10~m raster and the from 5~m to 10~m downsampled raster and the 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), start point to point3 (S-P3) and start point to point4 (S-P4).}
	\label{tab:6}
	\centering
	\begin{tabular}{ r @{\hspace*{3mm}}  r  r  r  r}
		Route & Method & $length /m$ & $costs_{al=f}$ & $d_{mean}$ /m \\
		\hline
		P1-E & 5~m 			& 107889.6 & 208547.8 &        \\
		 	 & Clipped 		& 107889.6 & 208547.8 &   0.0  \\
		 	 & Down			&  96754.2 & 212911.0 & 628.1 \\
		 	 & 10~m 		& 107232.9 & 203010.2 & 103.5 \\
		\hline
		P2-E & 5~m 			& 103706.4 & 155567.9 &        \\
		 	 & Clipped 		& 103706.4 & 155567.9 &   0.0 \\
		 	 & Down 	    &  92403.3 & 158238.6	& 639.9 \\
		 	 & 10~m 		& 104249.9 & 149899.7 & 177.7 \\
		\hline
		S-P3 & 5~m 			& 102187.1 & 34503.8 	&         \\
			 & Clipped 		&  90377.1 & 37926.1 	& 4465.4 \\
			 & Down 	    &  94125.6 & 37574.9 	&  742.4 \\
			 & 10~m 		& 102461.6 & 32446.0 	&   81.2 \\
		\hline
		S-P4 & 5~m 			& 96449.2 	& 33865.5 	&  		\\
		 	 & Clipped 		& 96449.2 	& 33865.5 	& 0.0 	\\
			 & Down 		& 87861.1 	& 36462.7 	& 796.4\\
			 & 10~m 		& 96739.5 	& 31899.3	& 83.5 \\

>>>>

\begin{table*}[t]
	\caption{Least Cost Paths as length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$) and the maximum minimum distance~($d_{max}$) and the agg. costs. From the agg. costs the differences of the agg. costs and the agg. costs per resolution are computed.} 
	\label{tab:2}
	\centering
	\begin{tabular}{ r  r  r  r  r  r  r  r  r  r}
		res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m & $d_{max}/m$ & agg.  $ cost_{al=f}$ & agg. $ cost_{al=t}$ &  $\Delta $ costs & agg. $costs_{al=f} \times m$ & agg. $costs_{al=t} \times m$ \\
		\hline
		5 	& 76136.3	& 78002.0 &  126.0 & 1065.0 & 18665.9 & 19616.8 & -850.00 & 93329.6 &  97584.8 \\
		10 	& 75430.1 	& 77936.6 &  277.9 & 1590.0 &  8931.2 &  9731.2 & -799.95 & 89312.5 &  97311.8 \\
		25 	& 75422.9 	& 78422.9 &  313.8 & 1621.2 &  3354.9 &  3872.7 & -517.78 & 83871.7 &  96816.4 \\
		50 	& 76135.0	& 70620.0 & 1140.0 & 4950.0 &  1409.0 &  2300.1 & -891.05 & 70451.2 & 115003.7 \\
		100 & 76283.8	& 74120.7 & 1946.4 & 6016.6 &   640.5 &  1572.3 & -931.70 & 64051.6 & 167226.8 \\

@@

As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (bi-linear) interpolation.

The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all touched False) shows, that only downsampling to a resolution of 10~m produces a path that is relatively close the high resolution paths (see table~\ref{tab:5}).

@@

Bi-linear downsampling the of the high resolution raster to a medium resolution did not result in any benefits compared to an original medium resolution raster.
The aggregated cost per resolution of the Least Cost Path from the downsampled raster is higher than the normal medium resolution 10~m raster.
In addition, the distance from these paths to the high resolution path  is greater than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
%%\todo{interpretation: Why is the downsampling result so different? Don't know.}

\begin{table}[ht!]
	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched False raster for the four control routes. For the reference path constructed from 5~m  and 10~m raster and from 5~m to 10~m downsampled raster and 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), starting point to point3 (S-P3) and starting point to point4 (S-P4).}
	\label{tab:6}
	\centering
	\begin{tabular}{ r @{\hspace*{3mm}}  r  r  r  r}
		Route & Method & $length /m$ & $costs_{al=f}$ & $d_{mean}$ /m \\
		\hline
		P1-E & 5~m 			& 107889.6 & 208547.8 &        \\
		 	 & Clipped 		& 107889.6 & 208547.8 &   0.0  \\
		 	 & Down			&  96754.2 & 212911.0 & 628.1 \\
		 	 & 10~m 		& 107232.9 & 203010.2 & 103.5 \\
		\hline
		P2-E & 5~m 			& 103706.4 & 155567.9 &        \\
		 	 & Clipped 		& 103706.4 & 155567.9 &   0.0 \\
		 	 & Down 	    &  92403.3 & 158238.6	& 639.9 \\
		 	 & 10~m 		& 104249.9 & 149899.7 & 177.7 \\
		\hline
		S-P3 & 5~m 			& 102187.1 & 34503.8 	&         \\
			 & Clipped 		&  90377.1 & 37926.1 	& 4465.4 \\
			 & Down 	    &  94125.6 & 37574.9 	&  742.4 \\
			 & 10~m 		& 102461.6 & 32446.0 	&   81.2 \\
		\hline
		S-P4 & 5~m 			& 96449.2 	& 33865.5 	&  		\\
		 	 & Clipped 		& 96449.2 	& 33865.5 	& 0.0 	\\
			 & Down 		& 87861.1 	& 36462.7 	& 796.4\\
			 & 10~m 		& 96739.5 	& 31899.3	& 83.5 \\
